Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-06 paper.tex
Edits: plots, annals of applied statistics format, tables, etc.

\documentclass[aoas, preprint]{imsart}
\RequirePackage{natbib}

@@

%\usepackage[section]{placeins}
\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
\usepackage{geometry}
\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}

@@

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Question is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
\# of Experts per Question & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
\# Forecasts given by each Expert on a Question & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
\# Questions participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
\end{tabular}
\caption{Five-number summaries of our real-world data.
%: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.
}
\label{DataStats}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
\label{ExpertiseTable}
\end{center}
\end{table}

@@

\begin{figure}[h!]
\vspace{-2em}
\centering
\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-3.5em}

@@

\subfigure{
\hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
\label{SyntheticSigma2}
}
\subfigure{
\hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
\label{SyntheticBeta}
}
\subfigure{
\hspace{-1.5em} \includegraphics{Figures/SyntheticK}
\label{SyntheticK}
}

@@

\subfigure{
\hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
\label{SyntheticSigma2b}
}
\subfigure{
\hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
\label{SyntheticBetab}
}
\subfigure{
\hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
\label{SyntheticKb}
}

@@

\begin{table}[htbp]
   \centering
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
&&&&\\
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
   \end{tabular}
   \caption{Brier Scores based on 10-fold cross-validation. \textit{Scores by Day} weighs a question by the number of days the question remained open. \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open. The bolded values indicate the lowest scores in each column. The values in the parenthesis represent standard errors in the scores. }
   \label{prediction}
\end{table}

@@

\begin{figure}[h!]
\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}

\subfigure[\scriptsize SDLM In-Sample]{
\hspace*{-3em}  \includegraphics{Figures/CalibrationNONE10}
\label{CalibrationNONEa}
}
\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationLOG10}
\label{CalibrationPLATTLOG10}
}
\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationBRI10}
\label{CalibrationPLATTBRI10}
}
\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationFULL10}
\hspace*{-2em} \label{CalibrationPLATTFULL10}
}

\subfigure[ \scriptsize SDLM Out-of-Sample]{
\hspace*{-3em} \includegraphics{Figures/CalibrationNONE}
\label{CalibrationNONEb}
}
\subfigure[ \scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Log}
\label{CalibrationSTC-Log}
}
\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Bri}
\label{CalibrationSTC-Bri}
}
\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics{Figures/CalibrationSO}
\hspace*{-2em}\label{CalibrationBSTC-Log}
}
\restoregeometry
\caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively. The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
}
\label{Calibration-Out}
\end{figure}

@@

\begin{figure}[!ht]
\vspace*{-1em} 
\includegraphics{Figures/BiasesBoxplots}
\label{BiasesLOGBoxplots}
\vspace*{-2em}

@@

\begin{figure}[h!]
\vspace{-1em}
\centering
\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-3.5em}

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics{Figures/ExamplePlot1b}
\label{Example1b}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics{Figures/ExamplePlot2b}
\label{Example2b}
}

@@

If the experts were asked to provide personal information during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts. For instance, this paper modeled the bias separately within each expertise group. This, however, is by no means restricted to the study of bias or its relation to self-reported expertise. Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts. Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions. This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.

>>>>

\documentclass[aoas, preprint]{imsart}

@@

%\usepackage[section]{placeins}
%\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
%\usepackage{geometry}
%\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
%\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
%\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}
\usepackage[OT1]{fontenc}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\endlocaldefs

@@

\begin{table*}
\caption{Five-number summaries of our real-world data.}
\label{DataStats}
\begin{tabular}{m{2in}	lllllll }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Question is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
\# of Experts per Question & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
\# Forecasts given by each Expert on a Question & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
\# Questions participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
\end{tabular}
\end{table*}

\begin{table*}
\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
\label{ExpertiseTable}
\begin{center}
\begin{tabular}{llllllll }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\end{center}
\end{table*}

@@

\begin{figure}[h!]
\vspace{-2em}
\centering
\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-3.5em}

@@

\subfigure{
\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
\label{SyntheticSigma2}
}
\subfigure{
\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
\label{SyntheticBeta}
}
\subfigure{
\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
\label{SyntheticK}
}

@@

\subfigure{
\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
\label{SyntheticSigma2b}
}
\subfigure{
\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
\label{SyntheticBetab}
}
\subfigure{
\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
\label{SyntheticKb}
}

@@

\begin{table*}
   \caption{Brier Scores based on 10-fold cross-validation. \textit{Scores by Day} weighs a question by the number of days the question remained open. \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open. The bolded values indicate the lowest scores in each column. The values in the parenthesis represent standard errors in the scores. }
   \label{prediction}
   \centering
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
      \hline
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
\hline
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
\hline
   \end{tabular}
\end{table*}

@@

\begin{figure}[h!]
%\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}

\subfigure[\scriptsize SDLM In-Sample]{
\hspace*{-2em}  \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE10}
\label{CalibrationNONEa}
}
\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationLOG10}
\label{CalibrationPLATTLOG10}
}
\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationBRI10}
\label{CalibrationPLATTBRI10}
}
\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ In-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationFULL10}
\hspace*{-2em} \label{CalibrationPLATTFULL10}
}

\subfigure[ \scriptsize SDLM Out-of-Sample]{
 \hspace*{-2em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE}
\label{CalibrationNONEb}
}
\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Log}
\label{CalibrationSTC-Log}
}
\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Bri}
\label{CalibrationSTC-Bri}
}
\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$ Out-of-Sample]{
\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationSO}
\hspace*{-2em}\label{CalibrationBSTC-Log}
}
%\restoregeometry
\caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively. The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
}
\label{Calibration-Out}
\end{figure}

@@

\begin{figure}[!ht]
\vspace*{-1em} 
%\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
\label{BiasesLOGBoxplots}
\vspace*{-2em}

@@

\begin{figure}[h!]
\vspace{-2em}
\centering
\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-3.5em}

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
\label{Example1b}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
\label{Example2b}
}

@@

If the experts were asked to provide personal information during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts. For instance, this paper modeled the bias separately within each expertise group. This is by no means restricted to the study of bias or its relation to self-reported expertise. Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts. Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions. This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.

%It may also be of interest to study hidden processes with other goals besides maximizing sharpness subject to calibration. This can be easily achieved by changing the optimization criterion in Equation (\ref{OSE}) to an appropriate function that meets the new goal.
